Developing a preprint culture in biology

### 2013-01-24 ms.tex
commit rev 1

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Developing a preprint culture in biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America
\\
\bf{5} Environmental Science, Policy, and Management. University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

@@

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, will help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These
initiatives are important to fully exploit the potential of open preprint
servers. Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original community-building
purposes.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics \cite{gin11}.
Preprints are also integral to the culture of other scientific fields.  Paul
Krugman noted that, in economics, the \emph{traditional model of submit, get
refereed, publish, and then people will read your work broke down a long time
ago. In fact, it had more or less fallen apart by the early 80s} \cite{kru12}.
In addition to a section in arXiv, economists have the RePEc (Research Papers in
Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) discuss the preprint policies of major publishers in biology; and
3) review the most popular preprint servers currently available.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

@@

Posting manuscripts as preprints also has the potential to result in higher
quality science by allowing prepublication feedback from a large pool of
reviewers. Prepublication reviews by a small network of colleagues are common in
the biological sciences and form an important part of the scientific process.
These ``friendly'' reviews increases the chances of errors being caught prior to
publication. Furthermore, the formal peer-review process as a whole is
critically over-loaded. As the number of active scientists increases and the
pressure to publish increases, it is becoming difficult for journals
to find reviewers \cite{hoc09}.  At the same time, rejection rates are high in
most journals \cite{aar08,roh09}, and when not invited to submit a revision,
authors must start the process over again at another journal. As a result,
initiatives to reduce time from submission to publication have emerged across
the scientific community. Rohr et al.  \cite{roh09} called for the recycling and
reuse of peer-reviews: by attaching previous reviews and detailed replies to a
new submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. A widespread use of
preprint servers can achieve the same goal of reducing the time spent in review.
With a rich enough community of scientists depositing preprints, and commenting
on them, the process of an open pre-review can become widespread and will
overall increase the quality of first submissions \cite{hoc12}.

@@

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).
Submitting to preprint servers has become more common in the past few years,
and the quantitative biology section in arXiv is experiencing the fastest
growth in submissions \cite{cal12}. Yet, the number of biology papers
submitted to preprint servers still represents only a small fraction of the
total research produced in biology.

@@

We briefly discuss the main options to submit preprints to open servers:
arXiv.org, figshare, and the upcoming PeerJ and F1000Research.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv provides a reliable citation system for all eprints and is especially
popular in high-energy physics. Physicist Paul Ginsparg created arXiv in 1991
for theoretical high-energy physicists to communicate preprints via email and
ftp, and soon thereafter adopted the newly created world-wide web
\cite{jackson2002preprints}.  arXiv now receives over 7 000 submissions per
month (\url{http://arxiv.org/show_monthly_submissions}) and divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to biologists.

One aspect of arXiv that differs from other options is that it has a moderation
system, which requires that papers must be categorized by an endorser.  At least
one author of a paper must be an endorser that has previously submitted a paper
or has received permission to submit to a particular category. arXiv is now
administered by the Cornell University Libraries, with funding coming from
voluntary pledges by academic institutions along with matching funds from the
Simons Foundation \cite{arxiv_future}. This approach to financing is one of the
numerous measures that arXiv takes to ensure that the repository will remain
permanently available and submissions will be readable.

@@

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness of
scientists with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. There is pressure on researchers to orient
their publication strategy in order to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication.  They all contribute somehow in a general slowing
down of scientific progress.  Research activities and the publication process
are drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, It might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study, its contribution to science, should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.

One of drawback of making publication easier is the proliferation of studies of
uneven quality. A trade-off between the intensity of the peer-review filtering
and the benefits to science has been hypothesized \cite{Aarssen2012}.  With
increasingly stringent peer reviewing, the quality of published papers and
easiness of finding discoveries might increase, at the cost of censorship, an
increased load on authors and reviewers, and greater delays for publication.
Preprints are simply bypassing this model for what we believe is the progress of
science: they speed up the dissemination of scientific discoveries, impede
censorship and put on reader's shoulders the responsibility to judge originality
and pertinence.

@@

We thank Carl Boettiger, Mark Hahnel, and Hedvig Nenz\'en for helpful comments
on an earlier version of this manuscript.

\section*{References}
% The bibtex filename
\bibliography{refs}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{map.eps}
\end{center}
\caption { {\bf It can take several months, and even a few years, before a submitted
paper is officially published and citable.} Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. }
\label{fig:map}
\end{figure}

>>>>

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{The Case for Preprints in Biology}
}
% Insert Author names, affiliations and corresponding author email.
\\
Philippe Desjardins-Proulx$^{1, 2,\ast}$, 
Ethan P. White$^{3}$, 
Joel J. Adamson$^{4}$,
Karthik Ram$^{5}$,
Timoth\'ee Poisot$^{1,2,6}$,
Dominique Gravel$^{1,2}$
\\
\bf{1} Theoretical Ecosystem Ecology laboratory, Universit\'e du Qu\'ebec \`a Rimouski, Quebec, Canada.
\\
\bf{2} Quebec Center for Biodiversity Science, McGill University, Quebec, Canada.
\\
\bf{3} Departement of Biology, Utah State University, Utah, United-States of America.
\\
\bf{4} Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill, North Carolina, United-States of America.
\\
\bf{5} Environmental Science, Policy, and Management, University of California, Berkeley, California, United-States of America.
\\
\bf{6} International Network for Next-Generation Ecology.
\\
$\ast$ E-mail: philippe.d.proulx@gmail.com
\end{flushleft}

@@

Open preprint servers offer a great opportunity for open science, especially if
the community embraces the idea of discussing preprints. Initiatives like
Haldane's Sieve (\href{http://haldanessieve.org/}{http://haldanessieve.org/}), a
new blog discussing arXiv papers in population genetics, can help make arXiv
attractive for scientists looking to promote their work \cite{lom12}. These
initiatives are important to fully exploit the potential of open preprint
servers. Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original community-building
purposes.

Preprints began to gain popularity 20 years ago with the advent of arXiv, an
open preprint server widely used in physics and mathematics \cite{gin11}.
Preprints are also integral to the culture of other scientific fields.  Paul
Krugman noted that, in economics, the \emph{traditional model of submit, get
refereed, publish, and then people will read your work broke down a long time
ago. In fact, it had more or less fallen apart by the early 80s} \cite{kru12}.
In addition to a section on arXiv, economists have the RePEc (Research Papers in
Economics) initiative, which aims to create an archive of working papers,
manuscripts, and book chapters.

Despite the success of this approach in other fields, most manuscripts in
biology are not posted to preprint servers and are therefore not seen by more
than a handful of other scientists prior to publication. In this article, we: 1)
highlight the advantages of open preprint servers for both scientists and
publishers; 2) discuss the preprint policies of major publishers in biology; and
3) review popular options to submit preprints.

@@

The first and most often discussed advantage of open preprints is
speed (Figure~\ref{fig:map}). The time between submission and the official
publication of a manuscript can be measured in months, sometimes in years. For
all this time, the research is known only to a select few: colleagues, editors,
reviewers. Thus, the science cannot be used, discussed, or reviewed by the wider
scientific community. In a recent blog post, C. Titus Brown noted how posting a
paper on arXiv quickly led to a citation (arXiv papers can be cited) and his
research was used by another researcher \cite{bro12}. The current system of
hiding manuscripts before acceptance pose problems for both scientists and
publishers. Manuscripts that are unknown cannot be used and thus take more time
to be cited. It has been shown that high-energy physics, with its high arXiv
submission rate, had the highest immediacy among physics and mathematics
\cite{pra05}. Immediacy measures how quickly articles are cited.

@@

Posting manuscripts as preprints also has the potential to improve the 
quality of science by allowing prepublication feedback from a large pool of 
reviewers. In our experience, prepublication reviews by a small network of 
colleagues are common in the biological sciences and form an important part 
of the scientific process. These ``friendly'' reviews increases the chances 
of errors being caught prior to publication. Furthermore, the formal 
peer-review process as a whole is critically over-loaded. As the number of 
active scientists increases and the pressure to publish increases, it is 
becoming difficult for journals to find reviewers \cite{hoc09}.  At the same 
time, rejection rates are high in most journals \cite{aar08,roh09}, and when 
not invited to submit a revision, authors must start the process over again 
at another journal. As a result, initiatives to reduce time from submission 
to publication have emerged across the scientific community. Rohr et al.  
\cite{roh09} called for the recycling and reuse of peer-reviews: by 
attaching previous reviews and detailed replies to a new submission, both 
the editor and the referees can gauge the work done on the manuscript, and 
perhaps evaluate it with less prejudice. A widespread use of preprint 
servers can achieve the same goal of reducing the time spent in review. With 
a rich enough community of scientists depositing preprints, and commenting 
on them, the process of an open pre-review can become widespread and will 
overall increase the quality of first submissions \cite{hoc12}.

@@

In contrast to other disciplines, the field of biology has effectively no
preprint culture, with the exception of small pockets of primarily highly
quantitative research (\emph{e.g.}, epidemiology, population genetics).  While
submitting to preprint servers has become more common in the past few years, the
number of biology papers submitted to preprint servers still represents only a
small fraction of the total research produced in biology (Fig.~\ref{arxiv}).

@@

We briefly discuss several options to submit preprints to open servers:
arXiv.org, figshare, and the upcoming PeerJ and F1000Research.

@@

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in some branches of mathematics and physics.
arXiv provides a reliable citation system for all eprints. Physicist Paul
Ginsparg created arXiv in 1991 for theoretical high-energy physicists to
communicate preprints via email and ftp, and soon thereafter adopted the newly
created world-wide web \cite{jackson2002preprints}.  arXiv now receives over 7
000 submissions per month (\url{http://arxiv.org/show_monthly_submissions}) and
divides its submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The quantitative biology
category includes subcategories for Populations and Evolution, Quantitative
Methods and other categories that may be of interest to biologists.

Unlike the other options presented here, arXiv has a system of moderators 
and endorsors. At least one author of a paper must be an endorser that has 
previously submitted a paper or has received permission to submit. 
Moderators have the power to change the classification of a paper. 
arXiv is now administered by the Cornell University Libraries, with funding 
coming from voluntary pledges by academic institutions along with matching 
funds from the Simons Foundation \cite{arxiv_future}. This approach to 
financing is one of the numerous measures that arXiv takes to ensure that 
the repository will remain permanently available and submissions will be 
readable.

@@

\subsection*{Other options}

Scientific publishing is more diversified than ever and there are many
alternative options to submit articles before formal publications. 
Social networks such as ResearchGate can be used to submit preprints
\cite{lin12}. Also, if GitHub pushes openness further by opening the writing process,
open notebooks go even further by opening the entire scientific process
\cite{san11}.

@@

The ongoing discussions on the publication process, peer-reviewing and
alternative publication models are all symptoms of the current uneasiness
with the ever growing obsession with bibliographic metrics such as
the impact factor \cite{Fisher2012}. Researchers are pressured to orient
their publication strategy to maximize their number of publications and
total citations. A well-known consequence is to submit manuscripts first to the
most prestigious journals, and then resubmit to ``lower level'' journals as they
are rejected. The numerous negative impacts of such behavior have been discussed
in depth \cite{hoc09} and include a long delay between the time a manuscript is
finished to its publication. Research activities and the publication process
are drifting away from their fundamental object, namely the diffusion of novel
scientific discoveries.

Developing a preprint culture in biology will not solve all problems with the
current publication process. However, it might significantly reduce its negative
consequences. The role of peer-reviewing is to judge the scientific quality of a
study. It is the first barrier against the fraudulent and poor quality science
susceptible to impede scientific progress. In practice, the peer-review system
is not only used to evaluate scientific quality but also to judge pertinence. On
the other hand, preprints are not filtered, neither for their quality nor their
pertinence. Widespread adoption of preprint servers has the potential to shift
the diffusion strategy: journals would remain important to validate
publications, but the relevance of a study should
only be judged by many more readers than the typical two-four anonymous
reviewers. With a shift in the diffusion strategy, the role of traditional
journals and their editors would be to showcase scientific discoveries for
specialized readership.

Making publication easier can lead to the proliferation of studies of uneven
quality. A trade-off between the intensity of the peer-review filtering and the
benefits to science has been hypothesized \cite{Aarssen2012}.  With increasingly
stringent peer reviewing, the quality of published papers can improve at the
cost of censorship, an increased load on authors and reviewers, and greater
delays for publication.  Preprints are simply bypassing this model for what we
believe is the progress of science: they speed up the dissemination of
scientific discoveries and put on reader's shoulders the responsibility to judge
originality and pertinence.

@@

We dedicate this article to Aaron H. Swartz (1986-2013).  We thank Carl
Boettiger, Mark Hahnel, and Hedvig Nenz\'en for helpful comments on an earlier
version of this manuscript.

\section*{References}
\bibliography{refs}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{map.eps}
\end{center}
\caption { {\bf It can take several months before a submitted paper is
officially published and citable.} Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are
given access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated
into current research. }
\label{fig:map}
\end{figure}

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{figure_2.png}
\end{center}
\caption { Submissions to the quantitative biology section lag behind
physics, mathematics, and computer science.
}
\label{fig:arxiv}
\end{figure}
